Towards human SuperEEG

### 2018-10-08 paper/main.tex

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[11pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

%\usepackage{scicite}
\usepackage[numbers, square, sort, compress]{natbib}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[left]{lineno}
\linenumbers
% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

@@

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

\newcommand{\taskrecon}{S1}
\newcommand{\perexpcorrmaps}{S2}
\newcommand{\perexptaskrecon}{S3}
\newcommand{\perexptaskreconseparated}{S4}

@@

\title{Towards human SuperEEG}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Lucy L. W. Owen$^{1}$, Andrew C. Heusser$^{1, 2}$, and Jeremy R. Manning$^{1\ast}$\\\\
$^{1}$Department of Psychological and Brain Sciences, Dartmouth College,\\
Hanover, NH 03755, USA\\
$^{2}$Akili Interactive,\\
Boston, MA 02110, USA}

% Include the date command, but leave its argument blank.

@@

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

@@

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

\begin{abstract}
  Human \textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by
    Robert J. Sawyer in his popular science fiction novel \textit{The
      Terminal Experiment}~\cite{Sawy95}} entails measuring ongoing
  neural activity with perfect precision and at arbitrarily high
  spatiotemporal resolution.  Although true SuperEEG is impossible
  using existing methods, here we present a model-based method for
  \textit{inferring} neural activity at millimeter-scale spatial
  resolutions and millisecond-scale temporal resolutions using
  standard human intracranial recordings.  Our approach assumes that
  different people's brains exhibit similar spatial correlations, and
  that (all else being equal) neural activity at nearby locations will
  tend to be similar.  One can then ask, for an arbitrary individual's
  brain: given recordings from a limited set of locations in that
  individual's brain, along with the observed spatial correlations in
  other people's recordings, what would recordings most likely have
  looked like at \textit{other} locations in that individual's
  brain?\\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

\section*{Introduction}
Modern human brain recording techniques are fraught with
compromise~\citep{SejnEtal14}.  Commonly used approaches include
functional magnetic resonance imaging (fMRI), scalp
electroencephalography (EEG), and magnetoencephalography (MEG).  For
each of these techniques, neuroscientists and electrophysiologists
must choose to optimize spatial resolution at the cost of temporal
resolution (e.g., as in fMRI) or temporal resolution at the cost of
spatial resolution (e.g., as in EEG and MEG).  A less widely used
approach (due to requiring work with neurosurgical patients) is to
record from electrodes implanted directly onto the cortical surface
(electrocorticography; ECoG) or into deep brain structures
(intracranial EEG; iEEG).  However, these intracranial approaches also
require compromise: the high temporal and spatial resolutions of
intracranial recordings comes at the cost of substantially reduced
brain coverage, since safety considerations limit the number of
electrodes one may implant in a given patient's brain.  Further, the
locations of implanted electrodes are determined by clinical, rather
than research, needs.

@@

Another approach to obtaining high spatial and temporal resolution
neural data has been to collect fMRI and EEG data simultaneously.
Simultaneous fMRI-EEG has the potential to balance the high spatial
resolution of fMRI with the high temporal resolution of scalp EEG,
thereby, in theory, providing the best of both worlds.  In practice,
however, the signal quality of both recordings suffers substantially
when the two techniques are applied simultaneously (e.g., see review
by~\cite{HustEtal12}).  In addition, the experimental designs that are
ideally suited to each technique individually are somewhat at odds.
For example, fMRI experiments typically lock stimulus presentation
events to the regularly spaced image acquisition time (TR), which
maximizes the number of post-stimulus samples.  By contrast, EEG
experiments typically employ jittered stimulus presentation times to
maximize the experimentalist's ability to distinguish electrical brain
activity from external noise sources such as from 60 Hz alternating
current power sources.

@@

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from XXX--XXX hours, and
includes data recorded roughly from when the patients woke up each
morning, to before they went to sleep at the end of each day.  In
addition to typical bed-ridden hospital patient activities (e.g.,
lying in bed, reading, watching television, using personal electronic
devices, listening to music, visiting with family and friends, etc.),
the patients also performed a variety of experimental cognitive tasks
throughout their day (primarily list-learning memory tasks).  For the
purposes of the Dataset 1 analyses presented here, we aggregated all
data across each recording session, ignoring the particular activities
or tasks the patients were performing at any given moment.  We used
Dataset 1 to develop and debug our main SuperEEG approach, and to
examine the extent to which SuperEEG might be able to generate
task-general predictions.  Dataset 2 comprised multi-session
recordings from XXX electrodes implanted in the brains of XXX epilepsy
patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  Whereas
Dataset 1 included recordings taken during a wide variety of
behaviors, Dataset 2 included recordings taken as each patient
performed each of two memory tasks: a random word list free recall
task and a categorized word list free recall task.  We used Dataset 2
to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.

@@

We first applied fourth order Butterworth notch filter to remove 60 Hz
($\pm$ .5 Hz) line noise from every recording (from every electrode).
Next, we downsampled the recordings (regardless of the original
samplerate) to 250 Hz.  (This downsampling step served to both
normalize for differences in sampling rates across patients and to
ease the computational burden of our subsequent analyses.)  We then
excluded any electrodes that showed putative epileptiform activity.
Specifically, we excluded from further analysis any electrode that
exhibited an average kurtosis of 10 or greater across all of that
patient's recording sessions.  We also excluded any patients with
fewer than 2 electrodes that passed this criteria, as the SuperEEG
algorithm requires measuring correlations between 2 or more electrodes
from each patient.  For Dataset 1, this yielded clean recordings from
4168 electrodes implanted throughout the brains of 67 patients
(Fig.~\ref{fig:methods}A); for Dataset 2, this yielded clean
recordings from 2975 electrodes from 24 patients.  Each individual
patient contributes electrodes from a limited set of brain locations,
which we localized in a common space~\citep[MNI152;][]{GrabEtal06}; an
example Dataset 1 patient's 54 electrodes that passed the predefined
kurtosis test are highlighted in black and red.

@@

A second mechanism whereby a given region $x$ can contribute to the
recording at $\eta$ is through anatomical connections between
structures near $x$ and $\eta$.  We use spatial correlations in the
data to estimate these anatomical connections~\citep{BeckEtal18}.  Let
$\bar{R}$ be the set of locations at which we wish to estimate local
field potentials, and let $R_{s}$ be set of locations at which we
observe local field potentials from patient $s$ (excluding the
electrodes that did not pass the kurtosis test described above). In
the analyses below we define $\bar{R} = \cup_{s=1}^S R_s$.  We can
calculate the expected inter-electrode correlation matrix for patient
$s$, where $C_{s,k}(i,j)$ is the correlation between the time series
of voltages for electrodes $i$ and $j$ from subject $s$ during session
$k$, using:

@@

\begin{align}
\hat{N}_{s}(x,y) & = { \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j)\cdot \mathrm{z}(\bar{C}_{s}(i,j))}\label{eqn:subj_corrmat_num}\\
 \hat{D}_{s}(x,y) & = \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i)
                    \cdot W(y,j). \label{eqn:subj_corrmat_den}\\
 \hat{C}_s &= \mathrm{r}\left( \frac{\hat{N}_{s}}{\hat{D}_{s}} \right)
             \label{eqn:subj_corrmat}.
\end{align}
After estimating the numerator ($\hat{N}_{s}$) and denominator
($\hat{D}_{s}$) placeholders for each $\hat{C}_{s}$, we aggregate these
estimates across patients to obtain a single expected full-brain
correlation matrix ($\hat{K}$; Fig.~\ref{fig:methods}D):
\begin{align}
 \hat{K} = \mathrm{r} \left(  \frac{\sum  \hat{N}_{s}}{\sum \hat{D}_{s}}\right).\label{eqn:corrmat}
\end{align}
Intuitively, the numerators capture the general structures of the
patient-specific estimates of full-brain correlations, and the
denominators account for which locations were near the implanted
electrodes in each patient.  To obtain $\hat{K}$, we compute a
weighted average across the estimated patient-specific full-brain
correlation matrices, where patients with observed electrodes near a
particular set of locations in $\hat{K}$ contribute more to the
estimate.

@@

Let $\alpha_s$ be the set of indices of patient $s$'s electrode locations in
$\bar{R}$, and let $\beta_s$ be the set of indices of all other
locations in $\bar{R}$. In other words, $\beta_s$ reflects the locations
in $\bar{R}$ where we did not observe a recording for patient $s$
(these are the recording locations we will want to fill in using SuperEEG).
We can sub-divide $\hat{K}$ as follows:

@@

Let $Y_{s,k,\alpha_s}$ be the number-of-timepoints ($T$) by
length($\alpha_s$) matrix of (observed) voltages from the electrodes in
$\alpha_s$ during session $k$ from patient $s$. Then we can estimate the
voltage from patient $s$'s $k^{th}$ session at the locations in
$\beta_s$ using~\cite{Rasm06}:

\begin{align}
\hat{Y}_{s,k,\beta_s} &= ((\hat{K}_{\beta_s,\alpha_s}\cdot\hat{K}_{\alpha_s,\alpha_s}^{-1})\cdot Y_{s,k,\alpha_s}^\mathrm{T})^\mathrm{T}.\label{eqn:reconstruction}
\end{align}
This equation is the foundation of the SuperEEG algorithm.  Whereas we
observe recordings only at the locations in $\alpha_s$,
Equation~\ref{eqn:reconstruction} allows us to estimate the recordings
at all locations in $\beta_s$, which we can define \textit{a priori}
to include any locations we wish, throughout the brain.  This yields
estimates of the time-varying voltages at \textit{every} location in
$\bar{R}$, provided that we define $\bar{R}$ in advance to include the
union of all of the locations in $\alpha_s$ and all of the locations
at which we wish to estimate recordings (i.e., a timeseries of
voltages).

We designed our approach to be agnostic to electrode impedances, as
electrodes that do not exist do not have impedances.  Therefore our
algorithm recovers voltages in standard deviation ($z$-scored) units
rather than attempting to recover absolute voltages. (This property
reflects the fact that $\hat{K}_{\beta_s, \alpha_s}$ and
$\hat{K}_{\alpha_s, \alpha_s}$ are correlation matrices rather than
covariance matrices.)  Also, note that
Equation~\ref{eqn:reconstruction} directly requires computing a $T$ by $T$
matrix, which can become computationally intractable when $T$ is very
large (e.g., for the patient highlighted in Fig.~\ref{fig:recon}, $T =
20458799$). However, because Equation~\ref{eqn:reconstruction} is time
invariant, we may compute $Y_{s,k,\beta_s}$ in a piecewise
manner by filling in $Y_{s,k,\beta_s}$ one row at a time (using the
corresponding samples from $Y_{s, k, \alpha_s}$).

@@

\section*{Results}
We used a cross-validation approach to test the accuracy with which
the SuperEEG algorithm reconstructs activity throughout the brain.
For each patient in turn, we estimated full-brain correlation matrices
(Eqn.~\ref{eqn:corrmat}) using data from all of the \textit{other}
patients. This step ensured that the data we were reconstructing could
not also be used to estimate the between-location correlations that
drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).  For that held-out
patient, for each of their electrodes in turn, we used
Equation~\ref{eqn:reconstruction} to reconstruct activity at the
held-out electrode location, using the correlation matrix trained on
all other patients' data at $\hat{K}$, and using activity recorded
from the other electrodes from the held-out patient as
$Y_{s, k, \alpha_s}$.  We then asked: how closely did each of the
SuperEEG-estimated recordings at those electrodes match the observed
recordings from those electrodes (i.e., how closely did the estimated
$\hat{Y}_{s, k, \beta_s}$ match the observed $Y_{s, k, \beta_s}$?).

To illustrate our approach, we first examine and individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (red), and its associated
reconstruction (blue), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = XXX, p = XXX$).
%Where the traces do not align, the predicted
%trace tends to be rounded towards zero, reflecting the zero-mean prior
%implicit in Equation~\ref{eqn:reconstruction} (paired $t$-test
%comparing absolute values of actual versus predicted voltages:
%$t(XXX) = XXX, p = XXX$).
Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example electrode
(correlation: $r = XXX, p = XXX$).
%Like in the shorter example trace
%shown in panel A, the reconstructed voltages throughout the full
%recording tended to be smaller in magnitude than the original
%recording ($t(XXX) = XXX, p = XXX$).
This example confirms that the SuperEEG algorithm recovers the recordings
from this single electrode well.  Next, we quantify the algorithm's
performance across the full dataset.

@@

For each held-out electrode, from each held-out patient in turn, we we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: XXX; $t$-test comparing mean of
distribution of $z$-transformed per-electrode correlation coefficients
to 0: $t(XXX) = XXX, p = XXX$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

>>>>

\documentclass[11pt]{article}
\usepackage[numbers, square, sort, compress]{natbib}
\usepackage{newpxtext,newpxmath}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[left]{lineno}
\linenumbers

@@

\newcommand{\taskrecon}{S1}
\newcommand{\perexpcorrmaps}{S2}
\newcommand{\perexptaskrecon}{S3}
\newcommand{\perexptaskreconseparated}{S4}
\newcommand{\suppstats}{S5}

@@

\title{Towards human SuperEEG} 
\author{
  Lucy L. W. Owen$^{1}$,
  Andrew C. Heusser$^{1, 2}$, and
  Jeremy R. Manning$^{1\ast}$\\\\
$^{1}$Department of Psychological and Brain Sciences, Dartmouth College,\\
Hanover, NH 03755, USA\\
$^{2}$Akili Interactive,\\
Boston, MA 02110, USA}

@@

@@

\baselineskip24pt
\maketitle

\begin{abstract}
  Human \textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by
    Robert J. Sawyer in his popular science fiction novel \textit{The
      Terminal Experiment}~\cite{Sawy95}} entails measuring ongoing
  neural activity with perfect precision and at arbitrarily high
  spatiotemporal resolution.  Although true SuperEEG is impossible
  using existing methods, here we present a model-based method for
  \textit{inferring} neural activity at millimeter-scale spatial
  resolutions and millisecond-scale temporal resolutions using
  standard human intracranial recordings.  Our approach assumes that
  different people's brains exhibit similar spatial correlations, and
  that (all else being equal) neural activity at nearby locations will
  tend to be similar.  One can then ask, for an arbitrary individual's
  brain: given recordings from a limited set of locations in that
  individual's brain, along with the observed spatial correlations
  learned from other people's recordings, how much can be inferred
  about ongoing activity at \textit{other} locations in that
  individual's brain?\\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

\section*{Introduction}
Modern human brain recording techniques are fraught with
compromise~\citep{SejnEtal14}.  Commonly used approaches include
functional magnetic resonance imaging (fMRI), scalp
electroencephalography (EEG), and magnetoencephalography (MEG).  For
each of these techniques, neuroscientists and electrophysiologists
must choose to optimize spatial resolution at the cost of temporal
resolution (e.g., as in fMRI) or temporal resolution at the cost of
spatial resolution (e.g., as in EEG and MEG).  A less widely used
approach (due to requiring work with neurosurgical patients) is to
record from electrodes implanted directly onto the cortical surface
(electrocorticography; ECoG) or into deep brain structures
(intracranial EEG; iEEG).  However, these intracranial approaches also
require compromise: the high spatiotemporal resolutions of
intracranial recordings comes at the cost of substantially reduced
brain coverage, since safety considerations limit the number of
electrodes one may implant in a given patient's brain.  Further, the
locations of implanted electrodes are determined by clinical, rather
than research, needs.

@@

Another approach to obtaining high spatiotemporal resolution
neural data has been to collect fMRI and EEG data simultaneously.
Simultaneous fMRI-EEG has the potential to balance the high spatial
resolution of fMRI with the high temporal resolution of scalp EEG,
thereby, in theory, providing the best of both worlds.  In practice,
however, the signal quality of both recordings suffers substantially
when the two techniques are applied simultaneously (e.g., see review
by~\cite{HustEtal12}).  In addition, the experimental designs that are
ideally suited to each technique individually are somewhat at odds.
For example, fMRI experiments often lock stimulus presentation
events to the regularly spaced image acquisition time (TR), which
maximizes the number of post-stimulus samples.  By contrast, EEG
experiments typically employ jittered stimulus presentation times to
maximize the experimentalist's ability to distinguish electrical brain
activity from external noise sources such as from 60 Hz alternating
current power sources.

@@

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from XXX--XXX hours
(Fig.~\suppstats A), and includes data recorded roughly from when the
patients woke up each morning, to before they went to sleep at the end
of each day.  In addition to typical bed-ridden hospital patient
activities (e.g., lying in bed, reading, watching television, using
personal electronic devices, listening to music, visiting with family
and friends, etc.), the patients also performed a variety of
experimental cognitive tasks throughout their day (primarily
list-learning memory tasks).  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording
session, ignoring the particular activities or tasks the patients were
performing at any given moment.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might
be able to generate task-general predictions.  Dataset 2 comprised
multi-session recordings from XXX electrodes implanted in the brains
of XXX epilepsy patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17,
  KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
  KuceEtal18}.  Each recording session lasted from XXX--XXX hours
(Fig.~\suppstats B).  Whereas Dataset 1 included recordings taken during a
wide variety of behaviors, Dataset 2 included recordings taken as each
patient performed each of two memory tasks: a random word list free
recall task and a categorized word list free recall task.  We used
Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional details on both datasets.

@@

We first applied fourth order Butterworth notch filter to remove 60 Hz
($\pm$ .5 Hz) line noise from every recording (from every electrode).
Next, we downsampled the recordings (regardless of the original
samplerate) to 250 Hz.  (This downsampling step served to both
normalize for differences in sampling rates across patients and to
ease the computational burden of our subsequent analyses.)  We then
excluded any electrodes that showed putative epileptiform activity.
Specifically, we excluded from further analysis any electrode that
exhibited an average kurtosis of 10 or greater across all of that
patient's recording sessions.  We also excluded any patients with
fewer than 2 electrodes that passed this criteria, as the SuperEEG
algorithm requires measuring correlations between 2 or more electrodes
from each patient.  For Dataset 1, this yielded clean recordings from
XXX electrodes implanted throughout the brains of XXX patients
(Fig.~\ref{fig:methods}A); for Dataset 2, this yielded clean
recordings from 2975 electrodes from 24 patients.  Each individual
patient contributes electrodes from a limited set of brain locations,
which we localized in a common space~\citep[MNI152;][]{GrabEtal06}; an
example Dataset 1 patient's 54 electrodes that passed the predefined
kurtosis test are highlighted in black and red.

@@

A second mechanism whereby a given region $x$ can contribute to the
recording at $\eta$ is through (direct or indirect) anatomical
connections between structures near $x$ and $\eta$.  We use temporal
correlations in the data to estimate these anatomical
connections~\citep{BeckEtal18}.  Let $\bar{R}$ be the set of locations
at which we wish to estimate local field potentials, and let
$R_{s} \subseteq \bar{R}$ be set of locations at which we observe local
field potentials from patient $s$ (excluding the electrodes that did
not pass the kurtosis test described above). In the analyses below we
define $\bar{R} = \cup_{s=1}^S R_s$.  We can calculate the expected
inter-electrode correlation matrix for patient $s$, where
$C_{s,k}(i,j)$ is the correlation between the time series of voltages
for electrodes $i$ and $j$ from subject $s$ during session $k$, using:

@@

\begin{align}
\hat{N}_{s}(x,y) & = { \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j)\cdot \mathrm{z}(\bar{C}_{s}(i,j))}\label{eqn:subj_corrmat_num}\\
 \hat{D}_{s}(x,y) & = \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i)
                    \cdot W(y,j). \label{eqn:subj_corrmat_den}\\
 \hat{C}_s &= \mathrm{r}\left( \frac{\hat{N}_{s}}{\hat{D}_{s}} \right)
             \label{eqn:subj_corrmat}.
\end{align}
After estimating the numerator ($\hat{N}_{s}$) and denominator
($\hat{D}_{s}$) placeholders for each $\hat{C}_{s}$, we aggregate these
estimates across the $S$ patients to obtain a single expected full-brain
correlation matrix ($\hat{K}$; Fig.~\ref{fig:methods}D):
\begin{align}
 \hat{K} = \mathrm{r} \left(  \frac{\sum_{s=1}^S  \hat{N}_{s}}{\sum_{s=1}^S \hat{D}_{s}}\right).\label{eqn:corrmat}
\end{align}
Intuitively, the numerators capture the general structures of the
patient-specific estimates of full-brain correlations, and the
denominators account for which locations were near the implanted
electrodes in each patient.  To obtain $\hat{K}$, we compute a
weighted average across the estimated patient-specific full-brain
correlation matrices, where patients with observed electrodes near a
particular set of locations in $\hat{K}$ contribute more to the
estimate.

@@

Let $\alpha_s$ be the set of indices of patient $s$'s electrode locations in
$\bar{R}$ (i.e., the locations in $R_s$), and let $\beta_s$ be the set of indices of all other
locations in $\bar{R}$. In other words, $\beta_s$ reflects the locations
in $\bar{R}$ where we did not observe a recording for patient $s$
(these are the recording locations we will want to fill in using SuperEEG).
We can sub-divide $\hat{K}$ as follows:

@@

Let $Y_{s,k,\alpha_s}$ be the number-of-timepoints ($T$) by
$\left|\alpha_s\right|$ matrix of (observed) voltages from the electrodes in
$\alpha_s$ during session $k$ from patient $s$. Then we can estimate the
voltage from patient $s$'s $k^{th}$ session at the locations in
$\beta_s$ using~\cite{Rasm06}:

\begin{align}
\hat{Y}_{s,k,\beta_s} &= ((\hat{K}_{\beta_s,\alpha_s}\cdot\hat{K}_{\alpha_s,\alpha_s}^{-1})\cdot Y_{s,k,\alpha_s}^\mathrm{T})^\mathrm{T}.\label{eqn:reconstruction}
\end{align}
This equation is the foundation of the SuperEEG algorithm.  Whereas we
observe recordings only at the locations indexed by $\alpha_s$,
Equation~\ref{eqn:reconstruction} allows us to estimate the recordings
at all locations indexed by $\beta_s$, which we can define \textit{a priori}
to include any locations we wish, throughout the brain.  This yields
estimates of the time-varying voltages at \textit{every} location in
$\bar{R}$, provided that we define $\bar{R}$ in advance to include the
union of all of the locations in $R_s$ and all of the locations
at which we wish to estimate recordings (i.e., a timeseries of
voltages).

We designed our approach to be agnostic to electrode impedances, as
electrodes that do not exist do not have impedances.  Therefore our
algorithm recovers voltages in standard deviation ($z$-scored) units
rather than attempting to recover absolute voltages. (This property
reflects the fact that $\hat{K}_{\beta_s, \alpha_s}$ and
$\hat{K}_{\alpha_s, \alpha_s}$ are correlation matrices rather than
covariance matrices.)  Also, note that
Equation~\ref{eqn:reconstruction} requires computing a $T$ by $T$
matrix, which can become computationally intractable when $T$ is very
large (e.g., for the patient highlighted in Fig.~\ref{fig:recon}, $T =
20458799$). However, because Equation~\ref{eqn:reconstruction} is time
invariant, we may compute $Y_{s,k,\beta_s}$ in a piecewise
manner by filling in $Y_{s,k,\beta_s}$ one row at a time (using the
corresponding samples from $Y_{s, k, \alpha_s}$).

@@

\section*{Results}
We used a cross-validation approach to test the accuracy with which
the SuperEEG algorithm reconstructs activity throughout the brain.
For each patient in turn, we estimated full-brain correlation matrices
(Eqn.~\ref{eqn:corrmat}) using data from all of the \textit{other}
patients. This step ensured that the data we were reconstructing could
not also be used to estimate the between-location correlations that
drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).  For that held-out
patient, for each of their electrodes in turn, we used
Equation~\ref{eqn:reconstruction} to reconstruct activity at the
held-out electrode location, using the correlation matrix learned from
all other patients' data as $\hat{K}$, and using activity recorded
from the other electrodes from the held-out patient as
$Y_{s, k, \alpha_s}$.  We then asked: how closely did each of the
SuperEEG-estimated recordings at those electrodes match the observed
recordings from those electrodes (i.e., how closely did the estimated
$\hat{Y}_{s, k, \beta_s}$ match the observed $Y_{s, k, \beta_s}$?).

To illustrate our approach, we first examine and individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (red), and its associated
reconstruction (blue), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = XXX, p = XXX$).
%Where the traces do not align, the predicted
%trace tends to be rounded towards zero, reflecting the zero-mean prior
%implicit in Equation~\ref{eqn:reconstruction} (paired $t$-test
%comparing absolute values of actual versus predicted voltages:
%$t(XXX) = XXX, p = XXX$).
Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example electrode
(correlation: $r = XXX, p = XXX$).
%Like in the shorter example trace
%shown in panel A, the reconstructed voltages throughout the full
%recording tended to be smaller in magnitude than the original
%recording ($t(XXX) = XXX, p = XXX$).
This example confirms that the SuperEEG algorithm recovers the
recordings from this single electrode well.  Next, we used this
general approach to quantify the algorithm's performance across the
full dataset.

@@

For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: XXX; $t$-test comparing mean of
distribution of $z$-transformed per-electrode correlation coefficients
to 0: $t(XXX) = XXX, p = XXX$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

As a stricter benchmark, we compared the quality of these
across-participant reconstructions (i.e., computed using a correlation
model learned from other patients' data) to reconstructions generated
using a correlation model trained using the in-patient's data.  In
other words, for this within-patient benchmark analysis we estimated
$\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in turn,
using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions
serve as an estimate of how well data from all of the other electrodes
from that single patient may be used to estimate held-out data from
the same patient.  This allows us to ask how much information about
the activity at a given electrode might be inferred through (a) volume
conductance or other sources of ``leakage'' from activity patterns
measured from the patient's other electrodes and (b) across-electrode
correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of
within-patient correlations was centered well above zero (mean: XXX;
$t$-test comparing mean of distribution of $z$-transformed
per-electrode correlation coefficients to 0: $t(XXX) = XXX, p = XXX$).
However, the across-patient correlations were substantially higher
($t$-test comparing average $z$-transformed within versus across
patient electrode correlations: $t(XXX) = XXX, p = XXX$).  This is an
especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all
data from the patient whose data is being reconstructed.  We repeated
each of these analyses on a second independent dataset and found
similar results (Fig.~\ref{fig:corrmap}B, D; within versus across
reconstruction accuracy: $t(23) = 6.93, p < 10^{-5}$). We also
replicated this result separately for each of the two experiments from
Dataset 2 (Fig.~\perexpcorrmaps).  This overall finding, that
reconstructions of held-out data using correlation models learned from
\textit{other} patient's data yield higher reconstruction accuracy
than correlation models learned from the patient whose data is being
reconstructed, has two important implications.  First, it implies that
distant electrodes provide additional predictive power to the data
reconstructions beyond the information contained solely in nearby
electrodes.  (This follows from the fact that each patient's
electrodes are implanted in a unique set of locations, so for any
given electrode the closest electrodes in the full dataset are likely
to come from the same patient.)  Second, it implies that the spatial
correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

%correlations by brain area, distribution of correlations
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/corrmap}
  \caption{\textbf{Reconstruction quality across all electrodes in two
      ECoG datasets.}  \textbf{A. Distributions of correlations
      between observed versus reconstructed activity by electrode, for
      Dataset 1.}  The across-patient distribution (black) reflects
    reconstruction accuracy (correlation) using a correlation model
    learned from all but one patient's data, and then applied to that
    held-out patient's data.  The within-patient distribution (gray)
    reflects performance using a correlation model learned from the same
    patient who contributed the to-be-reconstructed electrode.
    \textbf{B. Distributions of correlations for Dataset 2.}  This
    panel is in the same format as Panel A, but reflects results
    obtained from Dataset 2.  The histograms aggregate data across
    both Dataset 2 experiments; for results broken down by experiment
    see Figure~\perexptaskreconseparated. \textbf{C.--D.  Reconstruction
      performance by location.} Each dot reflects the location of a
    single implanted electrode from Dataset 1 (Panel C) or Dataset 2
    (Panel D).  The dot colors denote the average across-session
    correlation, using the across-patient correlation model, between
    the observed and reconstructed activity at the given electrode
    location.}
  \label{fig:corrmap}
\end{figure}

The recordings we analyzed from Dataset 1 comprised data collected as
the patients performed a variety of (largely idiosyncratic) tasks
throughout each day's recording session.  That we observed reliable
reconstruction accuracy across patients suggests that the spatial
correlations derived from the SuperEEG algorithm are, to some extent,
similar across tasks.  We tested this finding more explicitly using
Dataset 2.  In Dataset 2, the recordings were limited to times when
each patient was participating in each of two experiments (Experiment
1, a random-word list free recall task, and Experiment 2, a
categorized list free recall task).  We wondered whether a correlation
model learned from data from one experiment might yield good
predictions of data from the other experiment.  Further, we wondered
about the extent to which it might be beneficial or harmful to combine
data across tasks.
